	Commit 1/Month

[?] Much better strategy. Starting perhaps 2022-08-31.

	"Hymn" To X

0 more recounting "dreams" in public. Anyways it is good to use diplomatic, Politically Correct language so no more "cog in Kapital" or "brain organ in the Matrix" rather "human capital". X could deliver an oration on their siblings in some other neighbourhood which I might parse as Black Noise containing roughly 1 bit of information. I have empathy, sympathy, and compassion. Anyways we could then discuss housing markets and say the local housing market actually and interior design or perhaps the literal News content of The New York Times. The True True is that it really is all perceptive subjective mooshy woo experience and perhaps X could say my interest in Francois Chollet is OK and I dunno value judgements but X functions and "knows 5 trillion trillion trillion trillion trillion trillion trillion True facts and tons about legal corpi which is helpful, useful, and economically valuable to someone" so I dunno inferring Rationality or rather trying to reconstruct some time sequence of first order decisions and implicit ideas about long run utility function maximisation in an unironic way providing me with bits about their utility function. A very skilled player of the 4 Dimensional Chess Game Of A Long Life by my assessment.

	Coffee Heads

OK trivially one can only feel one's qualiae and some would react to my The Social Network .txt Notes with "who cares about that PEA AEP observation?" And someone else might contend "it reveals consciousness itself, being aware of that observation, as well as a certain mid latency memory functional operation being executed in his brain organ, and errrm being a coffee head". Explaining it to a sober high schooler kiddo... Limitless, "take off like a plane phew", My Type Of Party "A gift such as Lazar's apparently cannot really be explained, but it can be felt and enjoyed." Nah but I mean pills. Red, Blue, it is clear and represent clarity and lucidity and then an observer could observe my Output and agree or disagree. Jeff Bezos and I look objectively similar and also share the 0 alarm and perhaps 0 meetings before 10:00 habits. "Oh, my 'God' Jeff, I would love love love working 19 hours per day for your awesome firm. Get those Amazon trees trees suggesting optimised content for the users, optimising user inter faces, C codes, or whatever I am tasked with."

	Synaptic Pruning Literature

Interesting in the context of my .txt Output. ARL at the Summer Program On Applied Rationality And Cognition.

	Multi Armed Bandit Cases

Literature. Examine cases such as given that it is a uniformly random permutation of the n uniform distributions on [0,1], [0,2], ..., [0,n] what sort of information can be discovered in how many queries? To deduce the permutation? Expected runtime algorithm?

	"Human Legible" "Human Explainable" Argumentations Or Facts Sequences

Perhaps that Kaggle on rhetoric and argument is interesting. Maybe one can study comments sections and infer something about structures which people might at least instantly agree with in a sorta kinda System 1 way and classify these structures. I know that I once output a sorta kinda "boring neo liberal" line to a "socialist" which had to do with potentially un rousing sequence that could be converted into something like:

I remember getting my film developed at the South Austin CostCompany.
It took a lot of time and energy and the results are OK.
I now have a "multi trillion megapixel" camera on my iPhone SE.
Luxury consumer technology goods like the Apple MacBook Pro 16.2in monitours.
Therefore "friendly competition" is good in markets.

Or it was something like this I would now classify as extremely dreck, banal, lacking insight, and frankly nearly empty headed insofar as actual market structure and comprehension is concerned. Dunno about college kids, ignorant undergraduates, and their Not Even Wrong or Wrong poetry.

In terms of the Irrationality on display on Reddit it is the True True real accurate model of reality I mean people will bite on a sequence like:

X shot up a school.
X owned 10 bottles of toothpaste.
Therefore X is a so called "creepy weird male shmautist".
X's choice to own few assets is different and "disgusting, repulsive, outgroup to the maximum degree, gross".

	Fly By Rachel Yun Zhang

Further reflected.

	Rationality Writing Style

Interesting from Victoria Xia: "Victoria Xia is former math olympian turned software engineer. In her former life she was a 2-time Gold Medalist at each of the China Girl's Math Olympiad [2011, 2012] and the European Girl's Math Olympiad [2012, 2014]. She's most likely found out in nature, on the tennis court, or curled up on the couch with a book". I aspire to factual framings, legalese, and outputting .txt which would strike even one of these mathematicians who insist on outputting .txt on "philosophical zombies" as well formulated. Hoping to do the day by day actions which might lead to coworkers outputting .txt such as "I observe Lazar is calm, focused, disciplined, peaceful, and tersely communicates useful information to me when we interface on critical important tasks and topics". Some Rationalists seem to write that they understand human brain organs on a much deeper structural or predictive level than they really do due to reading biorxiv and Peven Stinker's and Daniel Dennett's bibliographies but now I perceive that even "line of thought" is not so well formulated and perhaps "interal monologue" is a little better. I aspire to being more technical so something like: "Nets Hawk Katz thought that applying X technique to Y task was a good idea" now strikes me as rather woo and "Nets Hawk Katz orated on the X technique and the Y task" is a little bit more descriptive, actual factual, and superior in describing an accurate model of reality and the situation to the readership. The verb "thought" is a rather deeply suspicious one in the English language as is the noun "idea" though the latter functions communicatively in .txt. Kiddos who study Lincoln-Douglas Debate or purport to have the technical skill of discussing psyche may suggest Nets Hawk Katz could potentially have been a mala fide actor with intent to deceive over a partially mutually exclusive economically beneficial ArXiV activity and was leading Lazar astray knowingly and willingly.

	Reflecting Further On Writing

Some would output .txt and assert that they do not know what a "trance, rush, thrill of low latency competitive programming mind sport For The Win esque" means. Many in my readership may output .txt and classify this string as the sort of Not Even Wrong poetry I ought to omit in future writings and .txt compositions. A lawyer may output .txt on the existence of some written records and server pings.

	The Task

Produce optimal output and signals. What the firm wants from you. Simulate a large set of humanity and their potential reactions to your .txt output and cease offending potential future coworkers.

	Facts

Human memory is faulty and one on ones or meetings are questionable qualitative ways to make decisions. Engaging in Not Even Wrong poetry, value judgement laden commentary, and so on and so on... the lesson of the Telephone game is that communication is very tricky and I want to really emphasise the facts in a crystal clear technical way so from now on e.g. IQ is a clear reference to a score measured on an actual IQ test, nothing more, nothing less, it is precisely that cognitive performance metric, and any inference is on the reader.

	No More "Qualiae"

It communicates very little information to the readership frankly and ought to not be used along with "psych" woo babble strings like "shmautism" when more technical communication is superior such as: person A obtained prescription pharmaceutical B at time C after an oral representation to doctor D.

	Hyper Predictable Coworker

I think being as predictible as possible and sticking very simply to the most useful, direct, work related topics is a good idea. Quiet, calm, cool, serious, disciplined, peaceful.

	"By Definition" And "Functionally Implies An Observer Might Output"

I might throw up some more unoffensive Parenting Notes drafts soon to give potential mates and others an idea for the sort of rhetoric I might unironically endorse in my home schooling with my kiddos. Mostly free from my esthetic and value judgements and perceptions as I am somewhat curious as to how my kiddos will act and perceive the structure of reality without the psychic energy of public schooling "community" e.g. focus on structures and True facts.

	Facts Over Feelings

I aspire to simply stating facts so that my writing will strike the reader as mostly neutral. No more "disgusting, repulsive, dreck, trash, garbage, etc.", instead "I observed that girl X chose to live in a home residential environment full of suboptimal stressors which cause relative cognitive dysfunction and worse reproductive outcomes".

	Puzzles

Perhaps I will try hard and yet never produce strong mathematical research or clear 100 citations. But some will still truly think my puzzles technical skill is itself impressive and I may one day soon have 10 or 20 1st Solves. I will not represent it as "trivial" for it is by definition elite and exclusive though it sometimes feels effortless, easy, and natural to me of course the rush, thrill, trance of the low latency For The Win esque mind sport.

	Background Prerequisite Reading Sneer [?]

I contend that to begin to comprehend the bare base minimum of Rationality Twitter requires seriously parsing one's way through the logical arguments of ~25 books which appear for free on the Library Genesis platform system and being capable of lowkey passing quizzes on their content. Many in the so called "community" have verifiably Failed to do that.

	Dating Mating Profile Draft

[Might cold email[s] then post on Art Of Problem Solving, Art Of Problem Solving Facebook Group, Astral Codex X]

I am 35 years old, retired, and live in a house named "IO". I want to reproduce soon and home school our children. I enjoy a simple lifestyle and like intelligent 23-26 year olds with low anxiety and high libido, especially STEM graduate students [>50 citations] and those with a strong background in maths competitions [Gold Medal, >= 11 on the AIME].

	Suicidality Follow Up Thought

Maybe co opting that public source of information leak and conscious or sub conscious someone could take the time to code up a script which will ping friends if a friend is listening to a lot of tracks classified as "suicidal".

	B⚠NNED 10 Years Later Reflections

Should keep this type of content to my Diary but humorous... 2012-08-01. Of course the background here is that Flosstradamus was apparently Banned by the Chicago Police Department:

"
BAD NEWS: so CPD is saying our set at West Fest almost caused a riot so they've had us removed from the @maddecent Block Party in Chicago.
11:52 AM · Jul 29, 2012 · Echofon
"

	Unipolar New World Order Government

I dunno it is quite easy for some of us in the internet commentariat to feel that if the world is soon under one nation state... offensive to state the hypothesis quite clearly from my own subjective vantage point one would suppose that 中国 is performing highly at producing members of the literati and cognitive elite in the upper echelons of the ivory towers. I now have very severe fundamental foundational irreconcilable axiomatic and axiological differences with most modern humans and normies so I really ought to keep my political and philosophical thought private. Especially with Judeo Christian invocations of "God, "Jesus", "brother", "sister", and gendered Latin and Greek nouns currently being out of fashion.

	Vegetarian Propaganda

Yeah might keep these drafts over in the relevant Projects folder but I think it would be interesting to try and track real values and proxies for sreen resolution, monitours, hours per day exposures, overall image files exposures, as well as general animal image files and pro vegetarian sentiment. Might also be Good to analyse the efficacy of like carnist memes and informational web logs with claims about "health" and "cognition" which strike me as well formulated and reasonable.

	Animal Memes

I am kind of interested in tracking over time the front page of Reddit and just what kind of exposure people are having to animal memes and image files. I know that I often see the so called "cute" ones and sometimes mommy mothers animals and like still think maybe some people dunno just how much affection healthy mothers be feeling towards their young and kin and offsprings.

	Corpus Vocabulary

I kind of want to measure how repetitive language usage is in my own corpus. Think that if one tried to correlate "IQ" or something you might find I was in fact very very very repetitive and prone to using the same words over and over whereas others in superior cognitive conditions and environments might display more peacock tail plumage in that domain. I mean I kind of want to like fucking spot check dudes like Taleb really try and see what the Right tail outliers are and like come up with some posterior based on an actual underlying memory model for the proposition that the author did in fact consult a massive book corpus of like Latin and French phrases in the construction of their own massive .txt output.

	Impending Suicide Death Detection

Seems like an important task someone ought to get around to doing better upon.

	1 Simple Trick To Boost Your IQ By X Points

Funny advertising which I was exposed to as a young child on the internet. I uh think it would be humorous on an Astral Codex X post on nootropics or biological hacking. I would suppose that my lifestyle to a generic undergraduate could in fact easily confer on the order of 15 points sorta kinda no lie directly measured after 12 months. Interesting numeric claim to make because it sort of gives a hand wavey upper bound estimate on how much of a drop or hit I will take to subjective wellbeing if I move into an office job position.

	Sorting

Could be nontrivial papers to compose in this domain about various algorithms under assumptions about just how sorted the input array is or is expected to be.

	Being Dumb, Stoopid, Ignorant, Always Hungry, Always Learning And Growing

Just because something is trite... interested in IBM and Wimbledon Artificial Intelligence and sports statistics datasets. Anyways it was just last night I finally fucking realised that not only is the literal tennis table structure very important for legs and visually, but it is often in regulation play Dark Blue with a Red floor of the precise size and dimensions. Dunno about how Bright White the ball is or background colours. Funny stuff, thinking about knees, upper legs, and sports like say women's singles badminton.

	London Personal Income Taxes

Maybe I ought to keep this one to a pseudonymous Crazy Lilac Rationality web log but one ought to consider Boris Johnson and like how to navigate anonymous Ethereum wallets such that one can defraud and commit tax evasion upon the United Kingdom England and United States Of Amerika government and retire younger somehow I mean... Mitt Romney don't pay no tax, I'm just tryna protect my stacks, Rick Ross done told me that, these ninjas tryna hold me back, lo/no! Consider costs of living, bank loans structures, IO timelines. One can post tax filings along with donations in public.

	Jane Street Puzzles And IBM Research Ponder This Release Notifications

	https://www.janestreet.com/puzzles/current-puzzle/

It would actually be super simple to write up a script to ping the server say 1/minute and ping me when a Delta occurs I mean when a change happens. On IBM when the 404 switches to a puzzle appearing. I am still fiending for that first 1st.

	Datasets On Neighbourhoods And Economics Outcomes, Mobility

Of course it is quite surprising to see that most from the hood, from the projects, from the Lost Creek do not in fact go on to make it to the super big buckeroonies so ha there is that. Reference class. Those dreams sometimes come True I guess.

	Quantitative Researcher Projects

Gee I am going to go view like 10000 projects and then make a few of my own here shortly I think in the Projects folder or Mid Brow Statistiques folder. Maybe try and literally do some image, uh, errm, figure out some way to contemplate colour, typefaces, font size, what portion of a webpage is text boxes, perhaps modify based upon some score produced from image files, throw some sentiment analysis from image files in the mix with text sentiment analysis in models for whatever from the News in the Trading contexts.

	Folders Of Artists' Portfolios Generator

Thinking about a prediction task of say Non Fungible Tokens artwork digital files sales price got me thinking again about copyright on photography and the lack of an easy way for me to download say relatively high resolution images of an entire portfolio sorted chronologically and with perhaps good meta too like name/title of piece, as well as perhaps sales history. In any case now with raw text from a website or visual image machine learning it should be easy to write up a script to do this and toss out duplicates... maybe maybe maybe dunno how web scrapers go with raw text from the website to exclude random pictures of like the artist themself but want art and maybe restrict to first 1000 pages of Google image search results. Or consult those other art datasets from sources and that Ching-Yun Hsu web log post who knows.

	Map Colouring Games

Fascinating literature.

	Particle And Anti Particle Annihilation

Interesting. Game Theoretic analysis of that troops stars application game on my to do list.

	Program To Push Folder To Github

Would be useful. For non Linux too via the desktop website platform system.

	Poker Player Live "Reading Skill" Analyses

I want to collect a large dataset of tracked live poker hands, and track just how outperformant players are when they largely deviate from a precomputed theoretical sort of "Game Theoretic Optimum" for live play in some reference class or on some other dataset. In particular the idea is that sometimes players make moves one would otherwise think of as kind of wonky donkey donk air headed but it looks on television like they actually had some sort of an edge from a sick live read so I want to know if Igor Kurganov e.g. is actually one of the Greatest Of All Time live "mentalists".

	How Many Subsets Of n-ominoes Permit Regular Planar Tilings?

Related to the 2022-06 IBM Research Ponder This task.

	"Autism" Slurring

Frankly it has been used by a ton of ignorant undergraduates who are out of touch with the medical reality of this real condition, and I know that some think it is OK or even uplifting to think of themselves as like "high functioning autism" but it has also Truly been used by many extremely severely unintelligent uneducated University Of Texas At Austin undergraduates e.g. in a pejorative way towards significantly more intelligent, educated, detail oriented, technical, and logically aware people, primarily men.

	Diary And News Critically Reading Private

People can feel free to query me. Only push high quality content in public. Clear, delete my Github history, and perhaps make some edits to remove spammy comments in Notes documents for future readership.

	Optimal Background Colour For Zoom From Bed

I think that the Lilac works out pretty well and emphasises the Lilac tones in my own face. It is bright but not too bright, and ought to leave a positive emotive tone and memory in the viewership's mind.

	A Drunk Man Will Find His Way Home

Not sure what sorts of papers have been written up on the topic of other sorts of random walks in Z2 and when they do have a P=1 of returning or where threshold values are.

	Automated And Application For Temperature Control

I am not sure how precisely to optimise in Builder with respect to daily curves, and say humidity perception tracking if humidity is not easy to tightly control at all times.

	John Isner Versus Nicolas Mahut

Off the dome interview style one might Fermi estimate say 100 years of Wimbledon and 128 - 1 = 127 matches per is 12700 matches and say like even if it was a coin flip then [4 choose 2]/[2^4] go to a 5th set for obtaining the score of 2-2 after 4 sets. Then [10 choose 5]/[2^10] will be 5-5 after 10 games and then even say if the probability of pairing off the relevant game is a 0.9 advantage there is something like an expected number of matches going this long or longer to be ~ 1.53 so this is a really interesting spot to try and come up with a posterior on say the match being rigged and the players being in Kahoots to form a world record. Got me thinking about the precise details of claims about the most unexpected sporting statistics across all sports and like random pointless events captured in the Guinness World Records book which I viewed a few times as a child. Bad model.

	Auction Rigging

Consult the literature on the game theory, credence, optimality of actions and strategies say if one is bidding on one's own auction to price manipulate in this market.

	Simple Typeface Ideas

Forget what has been done before, been a long time since I deeply studied these. Would love to do simple lines, simple lines restricted to vertical and horizontal, as well as a restriction to circle arcs and lines.

	Wikipedia Quiz

Perhaps based upon local, Amerikan page views count... uh somehow use Natural Language Processing or human generated test questions perhaps multiple choice. Could give a very quick immediate little data point for a proxy of "generally half educated" in a formalised numeric notion. Could be a fun little one, I am a little curious about my own potential performance here which I know would be quite a lot worse really than plenty of "130 IQ" Westlakers in fact.

	538 Riddler Troops Allocation Task

I am not sure if this task has been resolved in the Game Theory literature. In a setting, there exist n discrete territories say with values. And each player can allocate their troops accordingly with the winner, judged under some metric, of each territory winning all of that territory's value. Worth examining coalition structures.

	The Boring Project

Perhaps frankly literally anything will fly to get me in for Quantitative Researcher or Software Engineer at a top trading firm at this point. Could demonstrate a bunch of Object Oriented Programming and principles maybe cook up like some lame exploration game something with maybe history and countries and like new countries and like I dunno trading and currencies foreign exchange between countries.

	Trading Computer Science Software Engineering Projects

I have observed many projects listed on Resumes and CVs which apparently rose beyond the level of utterly boring into slightly unboring enough... perhaps I can in fact soon get to crack a lacking and cook up some outperformant C for those Joshi tasks I left hanging a few months back and/or some linear algebraic parallel library.

	Tetris Game Theory

It is an interesting enough game, not in the combo setting but rather in the traditional Marathon setting or perhaps 40 Line clear setting under some metric where the earlier the line clears... really considering analysing the 5x10 well case with and without a variety of different bag settings and look ahead settings. Think the canonical bag and 5 look ahead might be interesting. Immediately the sort of 2^50*... positions suggests that a naive True full dynamic programming enumeration of states and state based action set analysis will not work out and so this particular task may be well suited for a little case study in Reinforcement Learning. Perhaps even this could constitute a cool little project or even a mediocre paper submission on say the sort of precise hand waved value evaluations of the look ahead feature function in strategy. Just how useful is it really? I was flashing back to the Tetris Effect from my youth and just how funny the latency is when you primarily feel like your conscious attention is exclusively on the look ahead bar. Also, could consider metrics in fact related to DAS, Delayed Automatic Shift, so that hypothetical key strokes are penalised. Dunno about 4 wide i.e. a 4x10 well. Maybe I can smash this out today and generate some nice graphics and visualisations, .gif files, .mp4 files even make it all look real good custom colours minimalism and everything and then yay maybe Neural Information Processing Systems will even take this type of esthetic garbage about such a deeply canonical and fascinating game and hoi polloi will love it and go wild and think I am what is litty lit lit and interesting and then choose to hire me finally yay list it as a Paper/Project. Nah but seriously for real maybe this is the Key Idea I need to secure the bag as a Quantitative Researcher or Algorithms Developer or whatever. Ensure to include memory of the bag used thus far in the state as well as say the T-spin move and similar moves in the transition function or else the hardcore observers will think it was a too casual paper and model. Unclear if a 0 bag game is worth analysing too. Metaphor to my life mane playing 4 Dimensional chess 5 trillion moves in advance. Rook to pawn To G69 and I end my move... I will vanquish thou at a children's dating mating genes game...

	Tron Light Bike And Snake Slither Games

Worth examining. Maybe we can also like study different coalitions teams with different speeds or functions. Discrete Z2 setting. Not even quite sure what has been said enumeratively about number of distinct paths or sets of paths from which starting points in which underlying points sets or asymptotics and comparisons. Simple setting of square and corner starting point. In the vein of that Inernational Mathematics Olympiad hook problem task composition technique of just enumerating anything and seeing where it goes to flip it into a paper submission to like the Electronic Journal Of Combinatorics if it gets mid or high brow enough.

	Front End Design On News Feed Application

I was thinking about about like colour hexes and Conrad Bassett-Bouchard. Now a User Experience Design Manager at Google. And the various things he did on the web log of theCollins Coalition at CoCo Scrabble. And Woogles.io. I might apply to Google Fi in fact but anyways thinking back to my News Feed application I want it to look really good. And be simple maybe simple horizontal rectangles and you click on the thing and it has all the raw .txt pre loaded 0 image files and then when you click it the .txt expands and then any relevant image files come in but it needs to have 0 latency on the .txt components in the feed. And maybe a summary up top to give the user the option to exit after reading that summary. In any case still working on some fundamentals and ideations for that project. And I want custom colour options but especially White on Black and Black on White as pre set button options for users. Call it New News.txt OK sketch this out a little further. I think I will pitch this 1 to Y Combinator right now cook it up mostly pro bono it just fuck it great idea.

	Statistical Learning And Inference Final Project Brain Storming

Analog to advertising where they maybe generate like 100 advertisements and then test iteratively to produce the best 1 maybe somehow try and take articles body texts to produce optimised Headlines for a news firm e.g. Kind of interested in predicting say views count from the body, or views count from the Headlines based on a more complete Headlines dataset, maybe this could be useful, maybe predict plays on YouTube films, Likes on Tweets, plays on Spotify system platform or whatever... dunno if this is all seriously well studied. Maybe simply hop in the mix on one of these Kaggle Deep Learning Tensor Graphics Processing Units tasks. Maybe there exist some bored YouTube or Spotify users and I could help them with playlist generation tasks.

	Natural Language Processing Text File Productions

This is the thing about an application right like OK we can provide OK book summaries to people who want to query a machine for an OK book summary. We have some online human generated sources on those topics. They communicate something about some possibly superficial depths of meanings, who can be quite sure or measure what. But if I am finna go on a nice little pleasant date with a girl say on a hike in New Braunfels and I want to read up on New Braunfels well I can read a ton of text files in advance. But I would kind of like my machine to know the words I had already read and then use that corpus in a trained way to produce for me a nice little 5 page summary which I could skim immediately prior to the date to really impress the chick and entertain her with random facts about the area I most liked.

	Podcast: The More Partially Examined Life

Maybe just start a little podcast here try and become mildly wealthy and extremely famous just get in the money off of a microphone and my own style of rapping ranting about whatever it is I think is funny or interesting in the news with my own oration analyses on the interconnected network try and pop off like Milo or Ben. Maybe I can just do Slavoj Zizek style ranting over clips from cinema and do little like 300s films on YouTube if I get a good microphone.

	Dr. Zarathustra Brady Logic And Satisfiability Literature

There existed some interesting papers and open tasks here which are certainly worth contemplating deeply at some point in time. In the mix to stew and brew upon if possible, unclear how much more background reading will be required or helpful.

	Lazar Ilic News .txt Aggregator

At some point I will definitely need to write up my own little script, natural language processing, to have a good set of 1000 headlines, as well as some selected paragraphs from say the top 100 articles of the day or so roughly and might post the code and output somewhere on my own personal Github page. This one is going to be a smash hit at the implicit name your price to opt in to pay I mean hit me at my Venmo any time to contribute to the funding of this poor struggling natural language processing artist. Miniature links for post facto clicking after skimming. Frequency, quantity, mix of "the news hoi polloi need not want", and also learning from their clicks.

	Metaculus Performance Hacking

Of course one ought to solve the News .txt Aggregator task and practice on an alternate first, produce robust performance, and then hop in the mix on the primary "Lazar" account. Read and scrape the entire site and run a variety of analyses on other users and auto analyses as part of maxing out the centaur style updating and post facto machine learning based "fixing" and automated updating prior to optimal perhaps latest time commit.

	Deep Learning

One task I thought maybe search a lower resolution image version for the lower resolution version of the target such that we are properly prone to overly matching and then execute further matching analysis upstream in those search regions of the higher resolution image.

There exist pre processing algorithms which can be executed upon a dataset pointset based upon X alone which may indicate that executing a k nearest neighbours or any variant thresholded functionised k nearest neighbours algorithm will be weaker e.g. one may think a degenerate case of a bunch of sets of k+1 close dudes is bad.

	Light Bike Game

Generalisations, graphs, game theory, combinatorics?

	IBM Research Ponder This Task

Sometimes one tries a natural lattice mesh of points, say to test some function, or inequality, or to randomly execute as starting points for a gradient descent. But, is there some smaller set, leading to more computational ease, which nevertheless performs highly on relevant queries for relevant data sets?

I would think that in the distribution it would make sense at the very least to have some distributions polynomially and exponentially approaching convex hull vertices and boundaries along various paths.

	Kaggle Upvoted Comments Analysis

Write up some natural language processing and a script to ping the Kaggle server to automatically take me from 5 bronze comments to 50 and hit the "Expert" rating on the platform.

	Betting Mechanism

If it was not written down for circumstances without markets or simply private individuals choosing to engage in this. I proposed that 2 parties, say 1 proposes to the other that they bet on an upcoming binary event using this mechanism. Then they can discuss a magnitude e.g. a simple negotiation hopefully leaking not too much information for a constant c. Then the mechanism itself simply consists of the 2 people simultaneously writing down a and b and then the bet is given by c*f(a,b) now for example one might hope to satisfy a variety of desiderata. If one uses f(a,b) is the natural simple bet with bettings odds implied at (a+b)/2 : 1-(a+b)/2 then for example perhaps some agents can be incentivized to write down true underlying credences and have a bet wherein the EV capture prior to updating on the new information is something or other. Now I mean the actual matters are the matters but I wanted to propose something where they aren't using a bet following a friendly and revealing discussion rather the 2 agents are interested in a kind of a "fair" way to immediately attempt to capture EV off of eachother say they both just think they have an edge or whatever.

	Another Natural Language Processing Thought

Maybe to help editors, myself included, of Wikipedia a simple trawl to collect potential blue link insertions in to articles so that Wikipedia becomes even more connected of a graph help give the reader more easy use value.

	Introduction To Introductions

I feel that quite a few people would really benefit from a big survey perhaps even with estimates on information and hours like not just a compilation of 10000 Wikipedia articles to read in one single text file but rather a sequence of theorems and definitions to give people some ideas about what is out there.

In terms of topic categories as a starting point: rationality books, economics, biology, maths, computer science, cities, great men, politics, classics, religion, mythology, scientists, food, art, countries, philosophy, etc.

	Anthropics

I pinged Robin Hanson prior to Grabby Aliens with the thought that our existence implies there was not a light speed expansive takeover singularity in the history of our light cone and that in conjunction with the age of the universe and perhaps the numbers on abiogenesis credence one would deduce a very low posterior on the probability of abiogenesis producing such a singularity and of course it would then follow that we don't have much evidence of being a particularly unusual instance of abiogenesis and thus our probability of producing such a singularity may be lower than thought a priori. In terms of "reference class"es my perception is that a large number of philosophers and thinkers are totally wonky wacky and out of touch with reality. They don't understand the philosophy of the mind, tbe brain organ, human diversity in terms of cognitive research. And they really lack an understanding of logic and computer science 101. A lot of not even wrong, just poo which is not well mathematically formulated at all really and precisification reveals just how empty and inane some of their textual strings are really.

	.mobi, .epub, .pdf to .txt File Conversion

So there exist many techniques such as calling the replace all function with line breaks, nulls, "PARAGRAPH", more line breaks in order to rapidly convert in to a legible .txt file of a book. But making this more automated would help some people. For example to rapidly produce these table of contents to notes .tex template files I use [ctrl c] [ctrl v] and then I call like 52 instances of replace all with " a" to " A" and "-a" to "-A" and then I replace all new lines with 3 new lines. Native .epub and .mobi to .txt file conversion can be very very good and it is worth trying multiple source files until producing a 0 error conversion and then calling the replace all function on "'" to "'" and """ to """ as desired e.g. one can even call on "Mr" to "Mr." and "Mrs" to "Mrs." etc.