Bayesian Data Analysis

Doing Bayesian Data Analysis

The Bayesian Choice

Prior Distribution, Likelihood, Updating, Posterior Distribution

Recall Standard Distributions

Binomial, Normal, Poisson, Exponential, Discrete Uniform, Cauchy, Multinomial, Logistic Regression, Rounding, Poisson Regression, Normal Approximation, Hierarchical Beta/Binomial, Simple Random Sampling, Hierarchical Normal/Normal, Hierarchical Poisson/Gamma, Finite Mixture, Hierarchical Logistic Regression, Power-Transformed Normal, Student-t, Stratified Sampling, Cluster Sampling, Completely Randomized Experiments, Randomized Block And Latin Square Experiments, Truncation And Censoring, Capture-Recapture, Linear Regression, Linear Regression With Unequal Variances, Straight-Line Fitting With Errors In X And Y, Hierarchical Linear Regression, Hierarchical Model For Factorial Data, Generalized Linear Models, Probit Regression, Hierarchical Overdispersed Poisson Regression, Multinomial For Paired Comparisons, Loglinear For Contingency Tables, Negative Binomial, Beta-Binomial, Student-t Regression, Multivariate Regression, Hierarchical Multivariate Regression, Nonlinear Regression, Differential Equation Model, Hierarchical Regression With Unequal Variances

5 - Hierarchical Models

One can (with an implicit assumption about populations perhaps or measurements being related) introduce some hyperparameters and then be sure that all updates are done fully through all of the variables and their inter relations. Computation via simulation apparently.

6 - Model Checking And Improvement

"Do the model's deficiencies have a noticeable effect on the substantive inferences?" Graphical posterior predictive checks.

7 - Modeling Accounting For Data Collection

In general the analysis of sample surveys with unequal sampling probabilities can be difficult because of the large number of potential categories, each with its own model for the distribution of y given x and a parameter for the frequency of that category in the population. At this point, hierarchical models may be needed to obtain good results. Superpopulation parameters. For valid Bayesian inference we must condition on all observed data, which means we need the joint likelihood. Modeling accounting for data collection.

8 - Connections And Challenges

Bayesian, non Bayesian, Bayes estimate for loss function, posterior.

9 - General Advice

Construction of models to relate complicated data structures to scientific questions, checking the fit of such models, and investigating the sensitivity of conclusions to reasonable modeling assumptions. A willingness to use many parameters, hierarchical structuring of models, model checking, an emphasis on inference in the form of distributions, the use of simulation as the primary method of computation, the use of probability models as tools for understanding, the importance of conditioning on as much covariate information as possible in the analysis of data, the importance of designing studiest to have the property that inferences for estimands of interest will be robust to model assumptions.

10 - Overview Of Computation

General computing approach based on successive approximation.

11 - Posterior Simulation

The Gibbs sampler and the Metropolis-Hastings algorithm.

12 - Approximations Based On Posterior Modes

The exercises did not seem too tricky.

13 - Special Topics In Computation

Some maths here.

14 - Introduction To Regression Models



15 - Hierarchical Linear Models



16 - Generalized Linear Models



17 - Models For Robust Inference



18 - Mixture Models



19 - Multivariate Models



20 - Nonlinear Models



21 - Models For Missing Data



22 - Decision Analysis

