%-------------------------
% Rover Resume - Base Template
% Link: https://github.com/subidit/rover-resume
%
% Shows code for various formatting options for different resume sections.
% Education and Projects have single-line headers; while Experience uses double-line.
% Some formatting codes are kept inline; consider \newcommand{cmd}{def}.
% Excludes hyperref and icons for readability; MVP version.
% Explore other templates for more options.
% Mix and match as desired. Be consistent with headers and sub-headers.
%------------------------

\documentclass[11pt]{article} % fontsize 10pt/11pt/12pt

\usepackage[margin=1in, a4paper]{geometry}
\setcounter{secnumdepth}{0} % remove section numbering
\usepackage{titlesec}

% FORMATTING SECTIONS
% \bfseries adds bold back to Section Headers
\titleformat{\section}{\large\bfseries\uppercase}{}{}{}[\titlerule]

% \bfseries adds bold back to Subsections (Company/University)
\titleformat{\subsection}{\large\bfseries}{}{0pt}{}

% \bfseries adds bold back to Subsubsections (Job Titles)
\titleformat*{\subsubsection}{\large\bfseries}

\titlespacing{\subsection}{0pt}{*0}{*0}
\titlespacing{\subsubsection}{0pt}{*0}{*0}

\usepackage{enumitem}
% label={} keeps bullets removed, but indentation preserved
\setlist[itemize]{label={}, noitemsep,left=0pt .. \parindent}

% Restore bold labels for the Skills section (Technical/Athletics)
\setlist[description]{font=\bfseries, itemsep=0pt}

\pagestyle{empty} % remove page number
\pdfgentounicode=1

\begin{document}

\begin{center}
	\begin{minipage}{0.5\textwidth}
		{\Huge\bfseries
			Lazar Ilic % Bold Name
		} \\ \medskip
		%Resume Template % Title [optional]
	\end{minipage} \hfill
	\begin{minipage}{0.4\textwidth}
		\raggedleft
		LazarI2@Illinois.Edu \\
		GitHub.com/Lazar-Ilic
	\end{minipage}
\end{center}

\section{Experience}
%=================%
% Note: \normalfont is used after \hfill to un-bold the location/date
\subsection{High Frequency Trading Firm \hfill \normalfont Europe}
\subsubsection{Quantitative Research Intern \hfill \normalfont July 2025 - August 2025}
\begin{itemize}
	\item Generated profitable insights for mid-frequency algorithms by studying time series, examining predictors, and backtesting. Considered implementation shortfall due to market impact as well as funding rates and taker fees.
\end{itemize}

\subsection{Amazon \hfill \normalfont Boston, Massachusetts}
\subsubsection{Software Engineer \hfill \normalfont July 2023 - December 2023}
\bigskip

\section{Education}
%=================%
% Bold University, Normal Degree, Normal Date
\subsection{University of Illinois Urbana-Champaign $|$ {\normalfont Master of Computer Science} \hfill \normalfont 2024}
\begin{itemize}
  \item 3.63$|$4.00 Parallel Programming, Advanced Bayesian Modeling, Database Systems, Programming Languages And Compilers, Data Mining, Data Curation
\end{itemize}

\subsection{University of Texas at Austin $|$ {\normalfont Bachelor of Science of Mathematics} \hfill \normalfont 2022}
\begin{itemize}
  \item 3.77$|$4.00 Algorithms, Differential Equations, Microeconomics, Logic, Linear Algebra, Generalised Linear Models, Scientific Computation, Game Theory, Statistical Learning
\end{itemize}

\section{Research}
%=================
\subsection{Optimised Event-Driven Selling \hfill \normalfont 2025}
\begin{itemize}
  \item Discussed Kelly optimal strategies for planning to sell a financial instrument under various levels of information. Performed literature review and wrote example codes in Python and Jupyter Notebooks. Tied in real-world issues and mentioned various details and complexities which emerge in such settings.
\end{itemize}

\section{Awards}
%===============================%
\begin{enumerate}[label=\null, left=0pt..0pt, itemsep=0pt]
	\item Project Euler $\sim$50 Leaderboards In A Row + 3 Gold Medals - Outperformed AI \hfill 2021-2025
	\item Alibaba Global Mathematics Competition Top 800 - \$600,000 Prize Pool \hfill 2022-2024
	\item CSES 1st, CodeForces 2124, Meta Hacker Cup 319th, Putnam 87th, USAMO 43rd, USAPhO
\end{enumerate}

\section{Projects}
%=================
\subsection{Bridgewater Metaculus Contest $|$ {\normalfont Metaculus.com/accounts/profile/125313/} \hfill \normalfont 2025}
\begin{itemize}
  \item Won \$112 in Bridgewater's forecasting competition: cleaned data, used statistics, pandas, knowledge, and intuition. Computed parameters to automatically track the mixture of a relatively noninformative prior with market information including movement. Generated distributions during hidden periods as an exercise in data curation.
\end{itemize}

\section{Skills}
%===========================%
\begin{description}
	\item[Technical] LaTeX, C, C++, Python + Libraries, SQL, GRE 338 + 5.5
	\item[Athletics] 17:30 5K Run, 143 kg Touch And Go Bench Press
\end{description}

\end{document}